\subsection{BMV as a direct discriminator of actuality-sourced geometry}

Among the near-term tests of the framework, the Bose--Marletto--Vedral (BMV) protocol has a special status. It is not merely one more experimental arena in which RA might differ from standard expectations. It directly probes one of the framework's central architectural commitments: the separation between reversible possibility dynamics and ledger-writing actuality.

In the Engine of Becoming language, the relevant claim is simple. Step~2 carries amplitudes over candidate continuations. Step~4 selects one candidate for actualization. Step~5 updates the effective metric from the already-actualized causal ledger. The standard BMV entanglement mechanism requires branch-dependent geometry to be operative already at the Step~2 level. RA denies this. In the branch-coherent regime---where the positional alternatives of the masses remain coherent and no irreversible positional actualization occurs during the interaction interval---there is no branch-resolved effective metric. The metric remains sourced by the common actual ledger, so any gravitational phase is common-mode rather than entangling.

The resulting prediction is sharp:
\begin{quote}
\textbf{RA BMV prediction.} In an ideal BMV realization with nongravitational entangling channels removed, the purely gravitational entangling invariant vanishes,
\[
\Delta \phi_{\mathrm{grav}} = 0,
\]
and therefore the protocol yields a strict null result for gravity-mediated entanglement.
\end{quote}

This is stronger than the statement that gravity is merely classical. The claim is specifically that the effective geometry belongs to what has already been actualized, not to the coherent branch structure of what may yet happen. In that sense, RA permits superposition over future geometry-producing growth channels, but not superposition of the presently actual metric.

The falsification condition is correspondingly clean. If a BMV-type experiment, after removal of credible nongravitational channels, shows a positive gravity-mediated entanglement signal, then the Step~2 / Step~5 separation in present-form RA fails. If the signal is null, the result is consistent with the actuality-sourced geometry picture and strengthens one of the framework's most distinctive architectural claims.
