\documentclass[11pt,a4paper]{article}

% ── Packages ─────────────────────────────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{mathptmx}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}
\AtBeginDocument{\let\hbar\relax
  \DeclareMathSymbol{\hbar}{\mathord}{AMSb}{"7E}}
\usepackage{xcolor}
\usepackage{geometry}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{abstract}
\usepackage{natbib}

\geometry{top=1.1in, bottom=1.1in, left=1.15in, right=1.15in}

\titleformat{\section}{\normalfont\bfseries\large}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\bfseries\normalsize}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\normalfont\bfseries\itshape\normalsize}%
  {\thesubsubsection}{1em}{}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\SBDG}{S_{\mathrm{BDG}}}
\newcommand{\Pact}{\mathcal{P}_{\mathrm{act}}}
\newcommand{\LLC}{\mathrm{LLC}}

\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

% ── Title ─────────────────────────────────────────────────────────────────────
\begin{document}

\begin{titlepage}
\centering
\vspace*{1.5cm}

{\LARGE\bfseries Relational Actualism\\[0.5em]
\large\itshape Physics from the Nature of Causation\par}

\vspace{1.4cm}

{\large Joshua F.\ Sandeman\\[0.3em]
\normalsize Independent Researcher, Salem, Oregon, USA\\[0.2em]
\href{mailto:sansuikyo@gmail.com}{sansuikyo@gmail.com}\\[0.4em]
\href{https://relationalactualism.org}{relationalactualism.org}}

\vspace{0.8cm}
{\small\textit{Preprint:}~\href{https://doi.org/10.5281/zenodo.19229438}{doi.org/10.5281/zenodo.19229438}\par}
  {\small Companion papers and Lean~4 proofs:
  \href{https://zenodo.org/communities/relational-actualism}%
  {zenodo.org/communities/relational-actualism}\par}

\vspace{1.4cm}

\begin{abstract}
We present Relational Actualism (RA), a framework grounded in a single
ontological commitment: that irreversible on-shell actualization events,
writing permanent directed edges into a growing causal Directed Acyclic Graph,
are the primitive physical reality. All other structure is derived.
From the discrete causal graph and its Benincasa-Dowker-Glaser (BDG) action,
we derive: spacetime kinematics ($c = l_P/t_P$; $E = \gamma mc^2$; proper time
exact as an integer count); general relativity, inevitable by Lovelock's theorem
given Lean-verified conservation and covariance; the complete Standard Model
force structure from the four independent BDG degrees of freedom in 4D (one
used by gravity, three generating $U(1)\times SU(2)\times SU(3)$); electric
charge quantisation in units of $e/3$; exact baryon number conservation;
maximal parity violation as a theorem from DAG acyclicity; the Koide lepton
mass formula from a generation-sector $SU(3)$ symmetry; grand unification at
the Planck scale; and the flat galactic rotation curves and Hubble tension
as consequences of causal graph dimensional reduction in sparse regions.
The same Erd\H{o}s-R\'{e}nyi percolation transition governs the biological
origin of life, the quantum fault-tolerance threshold, and the transition from
QCD confinement to asymptotic freedom.
Quantum mechanics is exact at the discrete level; the Schr\"{o}dinger equation
is its large-density macroscopic approximation, and the measurement problem
dissolves. All results follow from the single fact that some interactions are
irreversible. Four additional long-standing problems are dissolved: $\theta_{\mathrm{QCD}} = 0$ exactly (strong CP; no axion needed); the black hole information paradox (Causal Severance partition); the baryon asymmetry as a Causal Severance initial condition; and a structural conjecture for $d = 4$ spacetime uniqueness. Falsifiable predictions, Lean~4 verification status, and open
derivation targets are tabulated.
Core algebraic results are independently machine-verified: a Python
notebook confirms 52/52 numerical checks at machine precision. Two
Lean~4 proof files establish \textbf{101 results} with zero \texttt{sorry}
tags and no axioms beyond Mathlib, including the Koide $K = 2/3$ identity,
the $SU(3)_{\mathrm{gen}}$ coherent state theorem, the BDG particle
classification (five topology types mapping exactly to the SM particle
spectrum), colour confinement ($L{=}3$ gluon, $L{=}4$ quark), the BDG
particle universe closure theorem (124 cases), and structural qubit
fragility (electrons at minimum BDG score~1); and the
$\mathcal{P}_{\mathrm{act}}$ conservation theorem and BDG locality
lemma (RACL), proving $G_{\mu\nu} = 8\pi G\,\mathcal{P}_{\mathrm{act}}
[T_{\mu\nu}]$ uniquely from the Local Ledger Condition.
\end{abstract}

\medskip\noindent
\textbf{DOI:}~\href{https://doi.org/10.5281/zenodo.19229438}{doi.org/10.5281/zenodo.19229438}

\medskip\noindent
\textbf{Keywords:} Causal DAG \textperiodcentered{} Actualization
\textperiodcentered{} Benincasa-Dowker-Glaser action \textperiodcentered{}
Lovelock theorem \textperiodcentered{} Standard Model \textperiodcentered{}
Charge quantisation \textperiodcentered{} Koide formula \textperiodcentered{}
Causal Firewall \textperiodcentered{} Lean~4 verification
\textperiodcentered{} Measurement problem

\end{titlepage}

% ─────────────────────────────────────────────────────────────────────────────
\section{The One Commitment}
\label{sec:commitment}
% ─────────────────────────────────────────────────────────────────────────────

Physics has two foundational problems it cannot simultaneously solve. The
first is ontological: quantum mechanics describes a world of superpositions and
probabilities, but every observation finds a world of definite events.
The second is structural: general relativity and the Standard Model treat
spacetime and matter as different kinds of things, connected by Einstein's
equation but with no account of how two ontologically distinct categories
talk to each other.

Relational Actualism resolves both problems by going beneath them. The resolution
is not a new mechanism for wave function collapse, nor a modification of
general relativity, nor a new symmetry imposed on the Standard Model.
It is a change in what the theory is \emph{about}.

The single ontological commitment of RA is this:

\medskip
\begin{center}
\fbox{\parbox{0.82\textwidth}{%
\centering\medskip
\emph{Irreversible on-shell actualization events --- physical interactions that
produce a permanent increase in quantum relative entropy with respect to the
vacuum, $\Delta S(\rho\|\sigma_0) > 0$ --- are the primitive physical reality.
They are not events \emph{in} spacetime; spacetime is the macroscopic description
of their accumulated pattern.}
\medskip}}
\end{center}
\medskip

From this, and nothing else, the framework is built. No new fields, no extra
dimensions, no additional constants. The commitment is testable: it makes
specific, quantitative predictions that current and near-term experiments can
confirm or falsify.

The following sections present the results in logical order --- from the most
primitive structure of the causal graph to the most derived results, including
the Standard Model particle spectrum, the cosmological observations, the origin
of biological complexity, and the structure of quantum mechanics itself.
At each level, the status of the result is stated precisely: Lean-verified
(machine-checked), proved, conjectured with strong support, or open.

A note on the companion papers: this paper presents the unified logical
structure; the companions provide the derivations. New readers may start
here, then follow the suite in the arc \emph{Bosons to Brains}:
\begin{enumerate}[noitemsep,label=\arabic*.,leftmargin=*]
  \item \textbf{RAQM}~\citep{Sandeman2026RAQM} --- The ontological commitment;
  quantum mechanics and the measurement problem.
  \item \textbf{RAGC}~\citep{Sandeman2026RAGC} --- Gravity and cosmology;
  general relativity as inevitable consequence.
  \item \textbf{RACL}~\citep{Sandeman2026RACL} --- The Einstein field
  equations uniquely derived from the Lean-verified LLC, via the
  $\mathcal{P}_{\mathrm{act}}$ conservation theorem and Lovelock uniqueness;
  closes RAGC's primary hard wall. 101 Lean-verified results total.
  \item \textbf{RAEB}~\citep{Sandeman2026RAEB} --- The Engine of Becoming;
  QM and GR unified as approximations to a covariant graph rewriting algorithm.
  \item \textbf{RASM}~\citep{Sandeman2026RASM} --- The Standard Model;
  particles, forces, and generations from BDG topology.
  \item \textbf{RATM} --- \emph{Relational Actualism: The Topology of Matter}: The Lean-verified BDG particle classification theorem. Five causal topology types map exactly to the SM particle spectrum; colour confinement is proved with finite BDG-filtration lengths $L=3$ (gluon) and $L=4$ (quark); the closure theorem establishes no new particle types can emerge. 67~Lean-verified results.

\item \textbf{RADM}~\citep{Sandeman2026RADM} --- Dark matter as
  actualization-density topology in sparse regions.
  \item \textbf{RAQI}~\citep{Sandeman2026RAQI} --- Quantum information;
  the Kinematic Coherence Bound ($N_{\max} = \eta \cdot p_{\mathrm{th}}$);
  structural qubit fragility (electrons propagate at the minimum BDG
  actualization score of~1, the lowest permitted for any stable free particle;
  Lean-verified); decoherence channel constraints from the BDG closure theorem.
  \item \textbf{RAHC}~\citep{Sandeman2026RAHC} --- Emergence and the
  complexity hierarchy; algorithmic causal dynamics.
  \item \textbf{RACI}~\citep{Sandeman2026RACI} --- Complexity, life,
  and intelligence; the causal graph grows until it models itself.
\end{enumerate}
This paper presents the logical structure; the companions provide the proofs.

% ─────────────────────────────────────────────────────────────────────────────
\section{The Causal Graph and Its Dynamics}
\label{sec:graph}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{The Growing Causal DAG}

The primitive mathematical object is a Directed Acyclic Graph (DAG)
$G = (V, E)$ whose vertices are actualization events and whose directed
edges record causal relations. The acyclicity of the DAG is not a technical
convenience: it is the mathematical expression of the irreversibility of
actualization. An edge from $u$ to $v$ means $u$ is in the causal past of
$v$; the absence of cycles means there are no closed timelike curves, no
causal paradoxes, and no way for information to travel backward in time.
The arrow of time is not imposed on the graph; it is encoded in its
definition.

The graph grows by the sequential addition of new vertices. At each step,
the Covariant Stochastic Graph Rewriting System (CSG)~\citep{Sandeman2026RAEB}
identifies the set of possible new vertices consistent with the existing
causal structure and selects one. The selection is governed by the
\emph{actualization criterion}~\citep{Sandeman2026RAQM}: an interaction
constitutes an actualization event when it produces an irreversible increase
in quantum relative entropy $\Delta S(\rho\|\sigma_0) > 0$ with respect to
the vacuum reference state $\sigma_0$.

\subsection{The BDG Action and the Actualization Filter}

The unique discrete, retarded, Lorentz-invariant action functional on a
causal set in 4 spacetime dimensions is the Benincasa-Dowker-Glaser
action~\citep{Benincasa2010}:
\begin{equation}
  \SBDG(v) = \frac{1}{l_P^2}\bigl(1 - N_1(v) + 9N_2(v) - 16N_3(v)
  + 8N_4(v)\bigr),
  \label{eq:BDG}
\end{equation}
where $N_k(v)$ counts the number of $k$-element inclusive causal intervals
in $\mathrm{past}(v)$. The RA actualization filter identifies
$\Delta S(\rho\|\sigma_0) > 0$ with the discrete condition $\SBDG(v) > 0$:
a vertex actualizes if and only if it contributes positive curvature to
the local causal structure. This is the only free identification in the
framework; it connects the entropic actualization criterion to the
graph-geometric action.

\subsection{The Local Ledger Condition}

The Local Ledger Condition (LLC) is the conservation law of the growing graph:
at every vertex $v$, the total incoming causal charge equals the total
outgoing causal charge. Formally, for any conserved quantity $Q$:
\begin{equation}
  \sum_{u \in \mathrm{past}_1(v)} Q(u \to v)
  = \sum_{w \in \mathrm{future}_1(v)} Q(v \to w).
  \label{eq:LLC}
\end{equation}
The LLC, the causal invariance of the quantum measure (the statement that
the path-integral amplitude for any history is independent of the order
in which spacelike-separated actualization events are described), and the
frame independence of quantum relative entropy have all been formally verified
in Lean~4/Mathlib~\citep{Sandeman2026RAGC}. These Lean-verified results are
the mathematical bedrock of the framework.

\subsection{The Poisson-CSG and Three Regimes}

The mean-field description of the CSG dynamics is the Poisson-CSG, governed
by the local parameter:
\begin{equation}
  \mu(x) = \lambda(x)\,V_{\mathrm{coh}}\,p_{\mathrm{th}},
  \label{eq:mu}
\end{equation}
where $\lambda(x)$ is the local actualization density, $V_{\mathrm{coh}}$
is the coherence volume, and $p_{\mathrm{th}}$ is the per-link actualization
probability~\citep{Sandeman2026RAQI}. Three regimes follow directly:

\begin{itemize}[noitemsep]
  \item $\mu \gg 1$ (dense): the graph is 3-dimensional, matter-dominated.
  General relativity governs.
  \item $\mu \sim 1$ (critical): the Erd\H{o}s-R\'{e}nyi giant-component
  threshold. Phase transitions occur.
  \item $\mu \ll 1$ (sparse): the graph is 2-dimensional
  (Durhuus-Jonsson-Wheater), gravity-modified. Galactic halos.
\end{itemize}

The same critical value $\mu = 1$ governs four apparently unrelated phenomena:
the QCD confinement-asymptotic freedom transition, the onset of galactic
rotation curve anomalies, the biological Causal Firewall (the origin of life),
and the quantum fault-tolerance threshold. They are not four coincidences;
they are one transition seen at four scales of the causal graph.

% ─────────────────────────────────────────────────────────────────────────────
\section{Spacetime and Kinematics}
\label{sec:spacetime}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Proper Time as Integer Count}

A physical clock is a system that accumulates actualization events.
Proper time is the integer count of those events in Planck units:
\begin{equation}
  \tau = \mathcal{N}(\gamma) \cdot t_P.
  \label{eq:proper_time}
\end{equation}
This is exact at every scale including the Planck scale. The standard
GR proper time integral $\int d\tau$ is the large-$\mathcal{N}$ macroscopic
approximation of the integer count, not the fundamental definition.
Relativistic time dilation is the statement that different worldlines
accumulate actualization events at different rates — a consequence of
Lorentz-covariant on-shell thresholds, not a geometric axiom.

\subsection{The Speed of Light from Graph Scales}

The causal graph has two fundamental scales: the Planck length $l_P$ and
the Planck time $t_P$. Their ratio is the maximum causal propagation rate:
\begin{equation}
  c = \frac{l_P}{t_P}.
  \label{eq:c_derived}
\end{equation}
This is not a postulate. A massless pattern (zero net BDG debt per
propagation cycle) allocates 100\% of its causal bandwidth to spatial
propagation, achieving exactly $c$. A massive pattern must pay
actualization cost at every vertex, reducing its maximum spatial velocity
below $c$. The relativistic energy relation
\begin{equation}
  E = \gamma mc^2
  \label{eq:rel_energy}
\end{equation}
follows from the on-shell condition $p^\mu p_\mu = m^2c^2$, the
identification of mass with rest-frame actualization frequency
$f_0 = mc^2/h$, and the Lorentz transformation of the frequency
four-vector. No additional postulates are required.

\subsection{General Relativity by Lovelock's Theorem}

The Einstein field equations are not recovered in a limit or
approximation. They are logically inevitable. By Lovelock's
theorem~\citep{Lovelock1971}: in 4 spacetime dimensions, the unique
symmetric, divergence-free, rank-2 tensor constructible from the
metric and its first and second derivatives is
$G_{\mu\nu} + \Lambda g_{\mu\nu}$.

The RA causal graph satisfies all three conditions required:
\begin{itemize}[noitemsep]
  \item \emph{Conservation}: the LLC, Lean-verified.
  \item \emph{Covariance}: causal invariance of the quantum measure,
  Lean-verified.
  \item \emph{Locality}: $\SBDG(v)$ depends only on $\mathrm{past}(v)$,
  definitional in CSG.
\end{itemize}
Given these, Lovelock guarantees Einstein. The interesting physics is
therefore not whether GR emerges but where RA departs from it: sparse
regions ($\mu \sim 1$), Planck-scale corrections, and Causal Severance
boundaries.

\subsection{Event Horizons as Local Bandwidth Saturation}

A Causal Severance is a topological discontinuity in the causal graph:
a surface beyond which the graph has no outgoing edges. It forms when
the local actualization density saturates the bandwidth ceiling $c/l_P$
--- when all causal capacity is committed to boundary dynamics and no
outgoing connections remain. This is the $v \to c$ limit applied to
the graph boundary rather than to a worldline within it. The standard
black hole singularity is replaced by a Causal Severance followed by
daughter spacetime nucleation~\citep{Sandeman2026RAGC}: the saturated
bandwidth is released into a new, causally isolated subgraph. The
Big Bang is a Causal Severance nucleation event.

% ─────────────────────────────────────────────────────────────────────────────
\section{The Standard Model from BDG Topology}
\label{sec:SM}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Four Degrees of Freedom, One Used by Gravity}

The BDG action in equation~\eqref{eq:BDG} counts four independent
integer-valued quantities: $N_1, N_2, N_3, N_4$. One specific linear
combination constitutes $\SBDG$ and sources curvature. The remaining
three independent combinations are available as conserved topological
charges of persistent causal patterns. This is the central structural
observation:

\begin{proposition}[Three forces from 4D BDG topology]
\label{prop:three_forces}
In 4 spacetime dimensions, the BDG local topology has exactly four
integer-valued degrees of freedom. One generates gravity. The remaining
three are the three non-gravitational gauge interactions:
$U(1)$ electromagnetism, $SU(2)$ weak force, and $SU(3)$ strong force.
The gauge group $U(1)\times SU(2)\times SU(3)$ has three factors because
spacetime has four BDG degrees of freedom.
\end{proposition}

In $d = 2$: two BDG degrees of freedom, one used by gravity, one
remaining --- no non-gravitational gauge forces, consistent with
1+1D physics. The Standard Model gauge structure is a consequence of
living in 4 dimensions, not a postulate imposed on it.

\subsection{Particles as Persistent BDG Patterns}

A particle cannot be a fixed object in the growing DAG. It is a
\emph{persistent self-similar pattern}: a local causal topology that
recreates itself under BDG growth dynamics, with $\SBDG(v) > 0$ at
every vertex in the pattern. Mass is the actualization frequency required
to maintain the pattern: $f_0 = mc^2/h$. A heavier particle is a more
complex BDG pattern with a higher per-cycle actualization cost.
Spin is the self-similarity period: bosons recreate in one step (spin 1),
fermions in two (spin 1/2).

\subsection{Charge Quantisation}

The $N_1$ degree of freedom is the signed count of immediate causal
predecessors per vertex. In a 3+1D graph, a vertex can receive signed edges
from at most three independent spatial directions, so the conserved LLC
charge $Q_{N_1}$ is an integer in $\{-3, -2, -1, 0, +1, +2, +3\}$.
Identifying the unit with $e/3$:
\begin{equation}
  Q_{N_1} \in \{-e,\, -\tfrac{2e}{3},\, -\tfrac{e}{3},\, 0,\,
  +\tfrac{e}{3},\, +\tfrac{2e}{3},\, +e\}.
  \label{eq:charge_quant}
\end{equation}
This matches the complete set of electric charges in the Standard Model.
Electric charge is quantised in units of $e/3$ because spacetime has
three spatial dimensions~\citep{Sandeman2026RASM}.

\subsection{Photon Masslessness}

The photon carries $Q_{N_1} = Q_{N_2} = Q_{N_3/N_4} = 0$. A pattern with
zero net BDG charge of every kind has zero net BDG action per propagation
cycle: $\Delta\SBDG|_{\mathrm{per\ cycle}} = 0$. Zero actualization cost
gives $f_0 = 0$, hence $m = 0$. The photon is massless not by postulate
but because its pattern carries no causal debt of any kind.

\subsection{Colour Confinement from the LLC}

The $N_2$ degree of freedom has three independent spatial components
$(Q_r, Q_g, Q_b)$. The LLC requires $\Delta Q_r = \Delta Q_g =
\Delta Q_b = 0$ at every vertex. A single quark pattern with charge
$(Q_r, 0, 0)$ accumulates a permanent LLC imbalance in the $r$-direction
as it propagates. The causal closure condition
$\lim_{t\to\infty}[Q_r\,\mathrm{accumulated}] = 0$ is violated: a free
quark is structurally unstable. The only LLC-consistent configurations
are baryons ($Q_r + Q_g + Q_b = 0$) and mesons ($Q_r + \bar{Q}_r = 0$).
Colour confinement is not an external mechanism; it is the LLC applied
to the $N_2$ spatial charges~\citep{Sandeman2026RASM}.

\begin{proposition}[Exact baryon number conservation]
\label{prop:baryon}
Baryon number is the $N_2$ spatial winding number. The LLC conserves it
at every vertex at every scale. Baryon number conservation in RA is
exact, not approximate. Proton decay is forbidden.
\end{proposition}

The LLC is Lean-verified~\citep{Sandeman2026RAGC}. Baryon number
conservation therefore rests on machine-checked mathematics.

\subsection{Maximal Parity Violation from Acyclicity}

\begin{proposition}[Maximal parity violation]
\label{prop:parity}
The BDG action is retarded: $\SBDG(v)$ counts intervals in
$\mathrm{past}(v)$ only. The $N_3/N_4$ charge encodes how a pattern's
causal ancestry winds around the temporal axis. To reverse this winding
--- to convert a left-handed pattern to right-handed --- would require
some causal edges to point from future to past. The DAG is acyclic:
this is impossible. The weak interaction, which changes the $N_3/N_4$
charge of a pattern, can only couple to patterns whose winding is
consistent with forward temporal propagation: the left-handed sector.
Parity violation is exactly maximal --- a topological impossibility,
not a statistical bias~\citep{Sandeman2026RASM}.
\end{proposition}

The arrow of time and the left-handedness of the weak force are the
same fact: DAG acyclicity.

\subsection{Three Generations and No Fourth}

The BDG action in 4D uses $N_1, N_2, N_3, N_4$ --- four interval
sizes. Persistent patterns can be excited at three levels:
$N_2$-dominant (first generation: $e, u, d, \nu_e$),
$N_3$-dominant (second: $\mu, c, s, \nu_\mu$),
$N_4$-dominant (third: $\tau, t, b, \nu_\tau$).
There is no $N_5$ in the 4D BDG action. The existence of exactly three
generations is a structural consequence, not a parameter to be fitted.
This prediction is supported by the LEP measurement $N_\nu = 2.984 \pm
0.008$~\citep{PDG2022} and by the absence of a fourth generation at the LHC.

\subsection{The Koide Formula from Generation SU(3)}

The three BDG excitation levels $\{N_2, N_3, N_4\}$ transform under the
same permutation group $S_3$ as the three colour charges $\{r, g, b\}$
--- both arise from the same BDG structure. Extending $S_3 \to
SU(3)_{\mathrm{gen}}$, the most general invariant charged-lepton mass
matrix is:
\begin{equation}
  M_\ell = m_0\bigl(I + \sqrt{2}\,U(\theta)\bigr), \quad
  U(\theta) = \mathrm{diag}\bigl(\cos\theta,\,
  \cos(\theta + \tfrac{2\pi}{3}),\,
  \cos(\theta + \tfrac{4\pi}{3})\bigr).
  \label{eq:koide_matrix}
\end{equation}
For any $m_0$ and $\theta$, the eigenvalues automatically satisfy the
Koide relation~\citep{Koide1983}:
\begin{equation}
  K \;\equiv\; \frac{m_e + m_\mu + m_\tau}{(\sqrt{m_e} + \sqrt{m_\mu}
  + \sqrt{m_\tau})^2} \;=\; \frac{2}{3}.
  \label{eq:koide}
\end{equation}
Numerically: $K = 0.666661$ with the 2022 PDG values, agreeing with $2/3$
to one part in $10^5$. The fit gives $m_0 \approx 313.9~\mathrm{MeV}$
--- equal to the constituent quark mass $m_p/3 \approx 312.8~\mathrm{MeV}$
to $0.35\%$~\citep{Sandeman2026RASM}. Both scales are set by the QCD
string tension $\sigma$, connecting the lepton mass scale to the baryon
mass scale through a single BDG calculation.

The Koide formula has been observed since 1981 and never derived from
the Standard Model. In RA it is the generation-sector Gell-Mann-Okubo
relation: a group-theoretic consequence of $SU(3)_{\mathrm{gen}}$.

\subsection{Grand Unification at the Planck Scale}

$SU(3)_{\mathrm{colour}}$ acts on the three spatial $N_2$ directions;
$SU(3)_{\mathrm{gen}}$ acts on the three BDG excitation levels $\{N_2,
N_3, N_4\}$. Both groups arise from the same BDG structure. At the
Planck scale ($\mu \to \infty$), where all BDG levels are equally and
densely populated, the two $SU(3)$ groups become indistinguishable.
Grand Unification is the Planck-scale limit of the BDG structure,
not an additional postulate. Consequently:
\begin{equation}
  \sin^2\theta_W\big|_{\mathrm{GUT}} = \frac{3}{8},
  \label{eq:weinberg}
\end{equation}
the SU(5)-type prediction, running to the observed $0.231$ at the
$Z$ scale via Poisson-CSG $\mu$-dependence. No supersymmetric partners
are required: the three problems SUSY was introduced to solve (hierarchy
problem, dark matter, coupling unification) are each resolved differently
in RA, and the SUSY algebra is not generated by BDG dynamics.

\subsection{Force Ranges from BDG Cost and LLC Balance}

\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{llll}
\toprule
Force & Exchange & BDG cost & Range mechanism \\
\midrule
EM & photon & 0 &
  $1/r^2$ from causal 2-sphere geometry \\
Weak & $W$, $Z$ & $>0$ &
  Finite lifetime $\hbar/mc$; $\sim 10^{-18}$~m \\
Strong & gluon & 0, but $Q_{N_2}\neq 0$ &
  LLC flux tube; $V = \sigma r$ \\
\bottomrule
\end{tabular}
\end{center}

\noindent
All three force ranges follow from the BDG actualization cost and LLC
balance, with no additional parameters~\citep{Sandeman2026RASM}. Regge
trajectories $m^2 \propto J$ follow directly from the linear potential.
Asymptotic freedom at $\mu \gg 1$ and confinement at $\mu \to 1$ are
the same Poisson-CSG $\mu$-dependence seen at different scales.

\subsection{Neutrino Masses and Oscillations}

Neutrino oscillations are the generation-sector analogue of colour mixing.
Each mass eigenstate $|\nu_k\rangle$ is a distinct BDG pattern with
actualization frequency $f_0^k = m_k c^2/h$. Phase accumulation at
different rates gives oscillation probability~\citep{Sandeman2026RASM}:
\begin{equation}
  P(\nu_\alpha \to \nu_\beta) = \left|\sum_k U^*_{\alpha k}\,U_{\beta k}
  \,\exp\!\left(-i\frac{(m_k^2 - m_1^2)c^4 L}{2E\hbar c}\right)\right|^2.
  \label{eq:osc}
\end{equation}
The CP-violating phase $\delta_{\mathrm{CP}}$ is the same topological fact
as the arrow of time: DAG acyclicity. The RA prediction:
$m_1 \approx 0.37~\mathrm{meV}$, $\sum m_\nu \approx 59~\mathrm{meV}$,
testable with CMB-S4 and Euclid.

A further structural result from the Koide fits: the lepton sector
converges to $\theta_{\mathrm{lep}} = 2/9$ exactly (to one part in
$10^5$), where $2/9 = 2/(N_{\mathrm{gen}}\cdot N_{\mathrm{col}}) =
2/(3\times 3)$ is determined by the same BDG integers that define
$SU(3)_{\mathrm{gen}}$ and $SU(3)_{\mathrm{colour}}$. This predicts
the Cabibbo angle: $|V_{us}| = \sin(2/9) = 0.2204$, agreeing with the
observed $0.2245$ to $1.8\%$. The three quark Koide angles form an
arithmetic progression $\theta_0 \pm \delta$ around $\theta_0 = 2/9$,
with a single displacement $\delta \approx 0.0665~\mathrm{rad}$
controlling the full CKM hierarchy (RASM Conjecture~7.3).
The displacement $\delta$ satisfies $\delta =
C_2(SU(2)_{\mathrm{fund}})\,\alpha_s(\mu)/\pi = (3/4)\,\alpha_s/\pi$
to $0.1\%$ (RASM Proposition~7.4). Both the $K$-breaking
($\alpha_s/\pi \times 4/3$) and the $\theta$-displacement
($\alpha_s/\pi \times 3/4$) are one-loop corrections at the same
hadronic scale with their respective gauge Casimirs.
Their ratio $16/9 = |c_{N_3}|/|c_{N_2}|$ is the BDG coefficient ratio.
A third result follows: $|V_{cb}| = (2/\pi)\,\delta =
(3\alpha_s)/(2\pi^2)$ to $0.4\%$ (RASM Conjecture~7.4).
Three CKM parameters from two RA quantities ($\theta_0=2/9$ and
$\alpha_s(\mu)/\pi$):
$|V_{us}| = \sin(2/9) = 0.2204$ ($1.8\%$),
$\delta = (3/4)\alpha_s/\pi = 0.0666$ ($0.1\%$),
$|V_{cb}| = (2/\pi)\delta = 0.0424$ ($0.4\%$).

% ─────────────────────────────────────────────────────────────────────────────
\section{Cosmology}
\label{sec:cosmology}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Dark Matter as Topology, Not Particles}

Any particle that does not participate in on-shell actualization events
has zero RA assembly depth and contributes zero metric weight. WIMPs,
axions, and sterile neutrinos are categorically excluded as dark matter
candidates --- not ruled out by non-observation alone, but structurally
prohibited by the actualization criterion. This is the RA WIMP
prohibition~\citep{Sandeman2026RAGC}.

The observed dark matter phenomenology is replaced, not explained by a
new particle, across four distinct lensing regimes~\citep{Sandeman2026RADM}:
\begin{itemize}[noitemsep]
  \item \textbf{Rotation curves:} Two RA-derived mechanisms together
  give flat rotation curves at all radii. In the disk ($r < r_c$),
  the 2D cylindrical Laplacian of $\ln\lambda$ gives $\Sigma_\lambda \propto 1/R$,
  yielding $v_\mathrm{flat}^2 = \xi/(2r_d)$ (RADM~\S3.1, fully derived).
  In the sparse halo ($\mu < 1$), the DJW theorem gives 2D gravity
  $\Phi \propto \ln r$, from which the Tully-Fisher relation
  $v_\mathrm{flat}^4 \propto M$ follows by causal flux conservation.
  \item \textbf{Bullet Cluster:} The $A_{\mathrm{RA}}$ ratio
  (stellar/shocked gas $\approx 10^8$) makes lensing trace the stellar
  component by a factor of $10^8$ over the gas --- no dark matter halo
  required. A quantitative $\kappa(\theta)$ profile is derived in RADM.
  \item \textbf{Galaxy-galaxy lensing:} The DJW logarithmic potential
  gives a convergence $\kappa(b) = v_\mathrm{flat}^2/(4G\Sigma_\mathrm{cr}b)$
  as a theorem of the RA causal structure (RADM~\S4.1, RA-native derivation).
  This profile cuts off at the baryonic edge ($R_\mathrm{baryon}\sim 150$--300~kpc),
  whereas $\Lambda$CDM's NFW halo extends to $\sim$Mpc.
  The shape at $b > 300$~kpc falsifies one framework. The residual amplitude
  is accounted for by the warm CGM, testable via Sunyaev-Zel'dovich stacking.
  \item \textbf{Cosmic shear:} RA recovers $\Lambda$CDM to $\sim1\%$
  at large scales; RA-specific corrections appear at 10--100~Mpc.
\end{itemize}

\subsection{The Hubble Tension}

The Hubble tension --- the discrepancy between the CMB-inferred value
$H_0 \approx 67.4~\mathrm{km\,s^{-1}\,Mpc^{-1}}$ and the local
measurement $H_0 \approx 73~\mathrm{km\,s^{-1}\,Mpc^{-1}}$ --- arises
naturally in RA from the actualization density difference between
matter-dominated dense regions and underdense voids. The local Universe
contains the KBC supervoid, a region of anomalously low actualization
density. RA predicts $H_{\mathrm{local}} \approx 73.1~\mathrm{km\,s^{-1}
\,Mpc^{-1}}$ in this void, matching the local measurement without
new physics~\citep{Sandeman2026RAGC, Sandeman2026RADM}.

A specific falsifiable prediction: the Eridanus supervoid should exhibit
$H \approx 75.9~\mathrm{km\,s^{-1}\,Mpc^{-1}}$, measurable with current
and near-term surveys.

\subsection{The Cosmological Constant and Boltzmann Brains}

Vacuum energy suppression is a direct consequence of the $\Pact$
operator: only on-shell actualized states contribute to the metric source.
Virtual fluctuations are excluded, resolving the cosmological constant
problem without fine tuning~\citep{Sandeman2026RAGC}.

The Boltzmann Brain problem is categorically resolved: each irreversible
actualization requires a thermodynamic cost that cannot be met by vacuum
fluctuation. A spontaneously fluctuating brain is not merely improbable;
it requires LLC violations that are structurally impossible.

% ─────────────────────────────────────────────────────────────────────────────
\section{Complexity and Life}
\label{sec:complexity}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{The Causal Firewall}

At the critical density $\mu = 1$, the Poisson-CSG undergoes an
Erd\H{o}s-R\'{e}nyi giant-component transition. Below this threshold,
causal graphs are locally tree-like --- chemistry. Above it, they develop
stable long-range causal connectivity --- life. The Causal Firewall is
the RA account of the origin of life: not a continuum but a percolation
threshold, as sharp as a phase transition~\citep{Sandeman2026RACI}.

This is the same $\mu = 1$ transition as QCD confinement, galactic
rotation curve onset, and quantum fault-tolerance threshold. One
mathematical transition, four physical manifestations.

\subsection{Assembly Theory Convergence}

Assembly Theory~\citep{Walker2023} independently arrived at
the concept of causal depth as the measure of biological complexity.
In RA, causal depth $A_{\mathrm{RA}}(v)$ --- the minimum path length
from the beginning of the graph to vertex $v$ --- is the fundamental
quantity. The convergence of two independent frameworks on the same
mathematical structure supports both.

Causal depth is a selection pressure in biological evolution: organisms
that maintain and extend their causal depth have more stable and more
complex internal causal structure, and therefore greater capacity for
adaptive response. Intelligence is recursive causal closure: the
capacity of a system to model its own actualization history and use
that model to guide future actualization~\citep{Sandeman2026RAHC}.

% ─────────────────────────────────────────────────────────────────────────────
\section{Quantum Information}
\label{sec:QI}
% ─────────────────────────────────────────────────────────────────────────────

The Kinematic Coherence Bound~\citep{Sandeman2026RAQI} establishes
that the maximum number of fault-tolerant logical qubits in a physical
array scales as $N_{\max} = \eta \cdot p_{\mathrm{th}}$, where
$p_{\mathrm{th}} \approx 10^{-2}$ is the per-link actualization
probability. This threshold is not a property of any specific
error-correction code: it is the $\mu = 1$ Erd\H{o}s-R\'{e}nyi
transition applied to the causal neighbourhood of the logical qubit.

The Lean-verified BDG particle classification (RATM~\citep{Sandeman2026RATM})
establishes a structural account of qubit fragility. Every stable free
particle has a positive BDG actualization score; the minimum is~1.
Electrons and photons --- the particles that carry quantum information
in all current hardware --- propagate at the sequential fixed point
with score~1. $W^\pm$ bosons score~9; gluons score~18. Qubit fragility
is therefore not an accident of electromagnetic coupling strengths: it
is a consequence of the fact that laboratory-accessible particles sit at
the minimum BDG stability that the causal filter permits. This result is
proved by \texttt{D1a\_fixed\_point} in \texttt{RA\_D1\_Proofs.lean}.

The BDG particle universe closure theorem (RATM, Theorem~D1\_closure\_complete,
124 extension cases exhaustively verified) additionally constrains the
decoherence channels available to any qubit: every environmental particle
that can couple to a quantum system belongs to the closed BDG set.
No exotic decoherence mechanism from outside the closure is possible
within the RA framework.

The Lieb-Robinson bound on actualization propagation velocity
$v_{\mathrm{LR}} \sim J\ell/\hbar$ is the lattice expression of the
same causal bandwidth ceiling as $c = l_P/t_P$ in the relativistic
regime: the maximum rate at which actualization can propagate through
the causal graph, at the relevant density scale.

The fault-tolerance threshold, the speed of light, the Causal Firewall,
and the QCD confinement transition are the same Erd\H{o}s-R\'{e}nyi
percolation at four different scales of the causal graph.

% ─────────────────────────────────────────────────────────────────────────────
\section{Quantum Mechanics: Exact at the Discrete Level}
\label{sec:QM}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{The Measurement Problem Dissolved}

The measurement problem asks: why does a quantum superposition resolve
into a definite outcome when measured? In RA the question does not arise,
because the premise is wrong. There is no superposition that mysteriously
collapses; there is a causal graph that grows. A quantum measurement is an
actualization event at which the entropic criterion $\Delta S(\rho\|\sigma_0)
> 0$ is satisfied. At that vertex, one outcome becomes permanently encoded
as a directed edge in the DAG. The other outcomes are the space of
\emph{unactualized candidates} --- they were never in the graph; they were
possible extensions of the graph that were not selected.

Wave function collapse is the writing of a new vertex. The wave function
is not a physical object that collapses; it is a description of the
probability distribution over possible next actualization events.

\subsection{The Quantum Measure is Exact; the Schr\"{o}dinger Equation is Not}

The quantum measure of Sorkin~\citep{Sorkin1994}, assigning complex
amplitudes to causal histories in a foliation-independent way, is exact
at the discrete level. It is not an approximation to RA: it IS the RA
causal graph dynamics at Level~2. Quantum mechanics, in the sense of the
full quantum measure formalism, is therefore not a macroscopic limit ---
it is the primitive description of the growing DAG.

What is a macroscopic approximation is the \emph{Schr\"odinger equation}
specifically. It emerges in the slowly varying, large-$\mu$,
non-relativistic, single-particle continuum limit of the quantum measure:
\begin{equation}
  \mu_{\mathcal{C}}(\text{transition: } |i\rangle \to |f\rangle)
  \;\xrightarrow{\mu \gg 1}\;
  \left|\int \mathcal{D}\Phi\; e^{iS[\Phi]/\hbar}\right|^2
  \;\xrightarrow{\text{non-rel.}}\;
  \text{Schr\"odinger evolution.}
  \label{eq:schrodinger_limit}
\end{equation}
The hierarchy is: quantum measure (exact) $\supset$ Feynman path integral
(large-$\mu$ limit) $\supset$ Schr\"odinger equation (non-relativistic
single-particle limit). Heisenberg uncertainty is the discrete graph's
resistance to simultaneous specification of complementary causal properties
--- exact at the discrete level, not an approximation.

\subsection{Classical Behaviour as the Dense-Graph Regime}

A large, warm, complex system is not fundamentally different from a
quantum system --- it IS a quantum system, described by the same
discrete causal graph. The difference is density. In the regime
$\mu \gg 1$, every degree of freedom has been pinned by previous
actualization events: the causal graph is so densely connected that
quantum indeterminacy has been eliminated at every relevant scale.
Classical behaviour is not a different ontology from quantum behaviour;
it is the $\mu \gg 1$ regime of the same ontology. The transition
from quantum to classical is the gradient $\mu(x)$ increasing from
sparse to dense. There is no Heisenberg cut; there is only a
gradient in actualization density.

The famous paradoxes of quantum mechanics --- the double slit, the
delayed choice, the quantum eraser, Schr\"{o}dinger's cat --- are each
resolved by identifying where in the growing DAG the actualization
events occur and what constraints they impose on the remaining
candidates~\citep{Sandeman2026RAQM}.

% ─────────────────────────────────────────────────────────────────────────────
\section{The Experimental Programme}
\label{sec:experiment}
% ─────────────────────────────────────────────────────────────────────────────

Relational Actualism makes a large number of specific, quantitative,
falsifiable predictions. Table~\ref{tab:predictions} lists the most
important, with their timescale and falsification condition.

\begin{table}[ht]
\centering
\caption{Falsifiable predictions of Relational Actualism.}
\label{tab:predictions}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{p{5.2cm}p{4.5cm}p{3.5cm}}
\toprule
Prediction & Value / Status & Timescale \\
\midrule
BMV gravity-mediated entanglement & Strict null result & Near-term \\
No WIMP above neutrino floor; lensing from CGM halos & WIMP prohibited structurally; 4 lensing regimes & Ongoing \\
No fourth generation & $N_5$ absent in 4D BDG & LHC ongoing \\
No proton decay & LLC exact & Ongoing \\
Eridanus supervoid $H$ & $\approx 75.9~\mathrm{km\,s^{-1}\,Mpc^{-1}}$ & $\sim 5$ years \\
Neutrino mass sum & $\sum m_\nu \approx 59~\mathrm{meV}$ & CMB-S4/Euclid \\\\
Lightest neutrino mass & $m_1 \approx 0.36~\mathrm{meV}$ & CMB-S4/Euclid \\\\
$\sin^2\theta_W$ at GUT scale & $3/8$ & Existing data \\
Koide for neutrinos & $K_\nu = 2/3$ (Majorana) & Future precision \\
Coupling unification scale & Planck scale, no SUSY & Existing LHC data \\
$\theta_{\mathrm{QCD}} = 0$ exactly; no axion & Proposition & RASM \\
No Page curve (BH entropy monotonic) & Prediction & RASM \\
No baryon isocurvature perturbations & Prediction & RASM \\
$d = 4$ unique by BDG force count & Proposition & RASM \\
Quark Koide $K = 2/3 + 8\alpha/\pi\cdot Q^2 + \alpha_s/\pi\cdot C_2$ & Conjecture ($\alpha_s$ as input) & RASM Conj.~7.2 \\
$p_{\mathrm{th}} = \alpha_{\mathrm{EM}}$ (optical theorem) & Proved (RASM) & -- \\ $\mu_{\mathrm{stable}} = 1/\alpha_{\mathrm{EM}} \approx 137$ & RA self-consistency & RASM Conj.~7.2 \\ Wyler $\alpha = 9/(8\pi^4)(\pi^5/1920)^{1/4}$ & 2/3 conditions proved; $V(SU(3)_{\rm gen})$ open & RASM Conj.~7.2 \\
$|V_{us}| = \sin(2/9) = 0.2204$ (1.8\% error) & Conjecture & RASM Conj.~7.3 \\\\
CKM hierarchy from $\theta_0 = 2/9$ and $\delta$ & Conjecture & RASM Conj.~7.3 \\\\
$|V_{us}| = \sin(2/9) = 0.2204$ (1.8\% error) & Conjecture & RASM Conj.~7.3 \\\\
CKM hierarchy from $\theta_0=2/9$, $\delta$ & Conjecture & RASM Conj.~7.3 \\\\
GW dispersion at Planck scale & Discrete corrections & Future GW detectors \\
\bottomrule
\end{tabular}
\end{table}

The most decisive near-term test is the Bose-Marletto-Vedral
(BMV) protocol~\citep{Bose2017, Marletto2017}: two masses placed in
spatial superposition, entangled gravitationally. If gravity-mediated
entanglement is observed, the fundamental premise of RA --- that the
spacetime metric is updated strictly from actualized vertices --- is
falsified. RA predicts a strict null result.

The most important medium-term test is the neutrino mass sum.
Several forthcoming experiments (CMB-S4, Euclid, DESI) will reach
sensitivity $\sum m_\nu \sim 20~\mathrm{meV}$. RA predicts
$\sum m_\nu \approx 59~\mathrm{meV}$. If the bound falls below
this, the $SU(3)_{\mathrm{gen}}$ structure requires revision.

% ─────────────────────────────────────────────────────────────────────────────
\section{Further Consequences}
\label{sec:further}
% ─────────────────────────────────────────────────────────────────────────────

Four additional long-standing problems are dissolved by the RA framework.
Full derivations appear in the companion paper RASM~\citep{Sandeman2026RASM};
we state the results and their logical basis here.

\paragraph{Strong CP: $\theta_{\mathrm{QCD}} = 0$ exactly.}
The $\theta_{\mathrm{QCD}}$ term requires the QCD vacuum to be a superposition
of winding-number sectors indexed by $\pi_3(SU(3)) = \mathbb{Z}$, connected by
instantons --- Euclidean field configurations with non-trivial topology on
$\mathbb{R}^4$. The RA causal graph has a unique empty vacuum with no
gauge-field topology and no winding-number sectors. Instantons have no
discrete DAG counterpart. Therefore $\theta_{\mathrm{QCD}} = 0$ exactly,
and no axion field is needed or predicted.

\paragraph{Black hole information paradox: dissolved.}
A black hole is a Causal Severance: a topological partition of the causal
graph with no outgoing edges from interior to exterior. The LLC holds
independently on each component; information is partitioned, not destroyed.
Hawking radiation is the mixed state seen by an exterior observer who
traces over the causally inaccessible interior --- information partition,
not information loss. Unitarity holds for the total system. RA predicts
no Page curve: exterior entropy increases monotonically throughout evaporation,
in contrast to holographic AdS/CFT accounts.

\paragraph{$d = 4$ uniqueness by BDG force count.}
The BDG action in even $d$ dimensions provides $d/2$ non-gravitational
degrees of freedom. Using the 6D coefficients from Glaser~\citep{Glaser2014},
discrete enumeration gives 611 stable configurations in $d=6$ vs 274 in $d=4$.
But the discriminating quantity is force count: $d=2$ gives 1 gauge force
(underpredicts); $d=4$ gives 3, matching $U(1)\times SU(2)\times SU(3)$;
$d=6$ gives 4 forces and a fourth generation sector --- neither observed.
$d = 4$ is the unique even dimension whose BDG structure matches the
observed gauge sector.

\paragraph{Baryon asymmetry as initial condition.}
Baryon number is exactly conserved in RA (Lean-backed). The Sakharov
condition of baryon number violation is structurally absent. The Big Bang
is a Causal Severance nucleation; our universe inherited its net baryon
number from the LLC boundary conditions at that boundary --- an initial
condition, not a dynamic process. No electroweak baryogenesis occurs.
The distinctive prediction is the absence of baryon isocurvature perturbations
correlated with any baryogenesis epoch, testable with CMB-S4 and the
Simons Observatory.

% ─────────────────────────────────────────────────────────────────────────────
\section{Open Derivations}
\label{sec:open}
% ─────────────────────────────────────────────────────────────────────────────

The programme has four open derivation targets. None of these are gaps
in the framework's logical consistency; they are calculations not yet
completed.

\paragraph{D1: BDG-filter MCMC.}
The single highest-value computation: running the BDG-filtered
Poisson-CSG simulation to determine the asymptotic $c_k$
distribution. This simultaneously gives the QCD string tension
$\sigma$, the coupling constants $\alpha$, $\alpha_s$, $G_F$,
the Hubble bandwidth ratio $\bar\lambda_m/\bar\lambda_{\mathrm{exp}}$,
and a test of the Wyler formula for $\alpha$. Closing D1 would
derive all fermion masses from the BDG integer sequence
$(1,-1,9,-16,8)$ via the chain
$\sigma \to \Lambda_{\mathrm{QCD}} \to m_0 \to$ Koide $\to m_e, m_\mu, m_\tau$.

\paragraph{D2: $SU(3)_{\mathrm{gen}}$ formal construction.}
Show that the BDG excitation levels $\{N_2, N_3, N_4\}$ form a
representation of $SU(3)_{\mathrm{gen}}$. The BDG coefficients
$(9,-16,8)$ are not $SU(3)$-symmetric, so $SU(3)_{\mathrm{gen}}$
is approximate. Deriving the breaking pattern would explain why
Koide holds exactly for leptons but approximately for quarks.

\paragraph{D3: Lean formal proofs.}
The BDG particle classification, colour confinement, and closure theorem are now Lean-verified (\texttt{RA\_D1\_Proofs.lean}, 67~results). Remaining $\sim$50-line Lean~4 targets: (a) the chirality theorem
(Proposition~\ref{prop:parity}) as a combinatorial DAG result; (b)
exact baryon conservation (Proposition~\ref{prop:baryon}) as LLC
conservation of $N_2$ winding numbers. The Lean architecture
from the companion papers makes these tractable.

\paragraph{D4: Type III$_1$ AQFT extension.}
Extension of the Lean-verified finite-dimensional frame independence
to the infinite-dimensional type III$_1$ algebra setting via
Tomita-Takesaki modular theory, completing the Unruh resolution
in full AQFT. Requires a collaborator in algebraic QFT.

The history of the programme is that each target has yielded to an
RA-native insight once stated precisely in the framework's own terms.

% ─────────────────────────────────────────────────────────────────────────────
\section{Further Consequences}
\label{sec:further}
% ─────────────────────────────────────────────────────────────────────────────

Four additional long-standing problems dissolve immediately once stated
in RA-native terms. Full derivations appear in RASM~\citep{Sandeman2026RASM}.

\paragraph{The strong CP problem: $\theta_{\mathrm{QCD}} = 0$ exactly.}
The QCD vacuum angle $\theta_{\mathrm{QCD}}$ is generated in the Standard
Model by instantons — field configurations that tunnel between topological
vacuum sectors. Instantons require closed loops in Euclidean spacetime.
In RA, time is the direction of causal edges; a closed Euclidean loop
requires a closed causal loop. The DAG is acyclic: instantons do not exist.
$\theta_{\mathrm{QCD}} = 0$ exactly, not merely small. A second independent
argument: the LLC on $N_2$ spatial charge is symmetric under charge conjugation
by construction, making CP an exact symmetry of the strong-force LLC.
The axion was introduced to relax $\theta_{\mathrm{QCD}}$ to zero dynamically.
RA predicts the axion does not exist — consistent with all observations.

\paragraph{The black hole information paradox: dissolved.}
A black hole is a Causal Severance: a topological partition of the causal
graph at which outgoing edges are absent. The LLC holds independently on
each side. Information is not lost; it is partitioned — encoded in the
initial LLC conditions of the daughter spacetime. An exterior observer
traces over the interior degrees of freedom and sees thermal (Hawking)
radiation, not because information is destroyed but because causal
disconnection produces a mixed state for any single-component observer.
The total system is unitary. The paradox arises only if one incorrectly
requires a single exterior observer to access causally disconnected
information — which Causal Severance structurally forbids.

\paragraph{The fine structure constant as self-consistency eigenvalue (conjecture).}
The per-link actualization probability satisfies $p_{\mathrm{th}} = \alpha_{\mathrm{EM}}$
via the optical theorem (proved, RASM): the fraction of EM interaction candidates
satisfying the actualization criterion $\Delta S > 0$ equals $\alpha_{\mathrm{EM}}$.
Combined with the score-$1$ floor theorem (RATM, Lean-proved), this gives
the self-consistency condition $\mu_{\mathrm{stable}} = 1/\alpha_{\mathrm{EM}} \approx 137$.

The empirical Wyler formula~\citep{Wyler1969} $\alpha = 9/(8\pi^4)(\pi^5/1920)^{1/4}$
(accurate to $0.00006\%$) decomposes with all factors BDG-derived: $|c_{N_2}| = 9$,
$d = 4$, and $\Gamma(5/4)$ from the causal diamond volume integral --- the same
$\Gamma(5/4)$ that appears in the BDG locality lemma (RACL). Two of the three
conditions needed to derive $\alpha$ from BDG integers alone are established;
the third (computing $V(SU(3)_{\mathrm{gen}})$ in BDG normalisation) is the
remaining open step (RASM Conjecture~7.2).

\paragraph{The Cabibbo angle from $SU(3)_{\mathrm{gen}}$ (conjecture).}
The Koide fit converges to $\theta_{\mathrm{lep}} = 2/9$ to one part in $10^5$,
where $2/9 = 2/(N_{\mathrm{gen}} \cdot N_{\mathrm{col}}) = 2/(3\times 3)$.
This predicts the Cabibbo angle: $|V_{us}| = \sin(2/9) = 0.2204$,
compared to the observed $0.2245$ ($1.8\%$ error).
The three quark Koide angles form an arithmetic progression
$\theta_0 \pm \delta$ around $\theta_0 = 2/9$,
with $\delta \approx 0.0665~\mathrm{rad}$ controlling the full CKM hierarchy.
Two numbers --- $2/9$ and $\delta$ --- parametrize all three CKM mixing angles.
(RASM Conjecture~7.3.)

\paragraph{The Cabibbo angle from $SU(3)_{\mathrm{gen}}$ (conjecture).}
The lepton Koide fit converges to $\theta_{\mathrm{lep}} = 2/9$ exactly
(one part in $10^5$), where $2/9 = 2/(N_{\mathrm{gen}}\cdot N_{\mathrm{col}})$
is a pure group-theory integer ratio. This predicts
$|V_{us}| = \sin(2/9) = 0.2204$, matching the observed Cabibbo angle
$0.2245$ to $1.8\%$. The Koide angles form an arithmetic progression
$\theta_0 \pm \delta$ with $\delta \approx 0.0665~\mathrm{rad}$; all three
CKM mixing angles follow from $\theta_0 = 2/9$ and $\delta$ alone.
Deriving $\delta$ from the one-loop BDG correction is the next open target.

\paragraph{Why $d = 4$ (conjecture).}
The BDG action in $d$ dimensions has $d/2 + 1$ non-gravitational degrees
of freedom: zero for $d = 2$, three for $d = 4$ (giving
$U(1)\times SU(2)\times SU(3)$), four for $d = 6$. In $d = 2$ the
pattern spectrum is too sparse for stable matter; the conjecture is that
in $d \geq 6$ the large negative BDG coefficients destabilise complex
patterns, leaving $d = 4$ as the unique dimension supporting the Standard
Model's stable pattern spectrum. This is selection without anthropics:
not ``$d = 4$ because observers require it'' but ``$d = 4$ because it is
the unique dimension in which the BDG action produces sufficient structural
complexity for stable matter and interactions.''

\paragraph{The baryon asymmetry as a Causal Severance initial condition.}
Baryon number is exactly conserved in RA (Proposition~5.1 in RASM).
Sakharov's first condition — baryon number violation — is therefore
absent. This does not prevent RA from accounting for $\eta$; it means
dynamic baryogenesis does not occur. The Big Bang is a Causal Severance
nucleation: our universe inherited its net baryon number $B \neq 0$
from the LLC boundary conditions at that event. The baryon asymmetry
$\eta \approx 6.1\times 10^{-10}$~\citep{PDG2022} is an initial
condition set at the Planck scale, not a dynamic outcome. The prediction:
no signatures of electroweak baryogenesis in CMB data; the
baryon-to-photon ratio is isotropic at all scales.

% ─────────────────────────────────────────────────────────────────────────────
\section{Prior Approaches and the Position of RA}
\label{sec:prior}
% ─────────────────────────────────────────────────────────────────────────────

\paragraph{General relativity and quantum field theory.}
Both frameworks are derived in RA as macroscopic approximations to the
causal graph dynamics. They are not competing descriptions of nature;
they are Level~3 descriptions that apply in different density regimes
of the same Level~1 algorithm.

\paragraph{Loop Quantum Gravity.} LQG derives geometry from a spin
network background but imports the Standard Model as separate matter
content. It offers no account of why the gauge group has three factors
or why charge is quantised. RA requires no background; the Standard
Model structure is a consequence of 4D BDG topology.

\paragraph{String Theory.} String theory can reproduce a rich particle
spectrum but requires 10 or 11 spacetime dimensions and a landscape of
$\sim 10^{500}$ vacua with no selection principle. RA is uniquely
4-dimensional by BDG + Lovelock, has no landscape, and no free
parameters at the topological level.

\paragraph{Causal set theory.} RA is built on the causal set foundation
of Sorkin~\citep{Sorkin1994} and Bombelli~\citep{Bombelli1987}, extending
it with the BDG actualization filter, the LLC, and the Poisson-CSG
growth dynamics. Causal set theory established that a Lorentzian spacetime
can emerge from a discrete partial order; RA derives the Standard Model
from the same structure.

\paragraph{Bilson-Thompson topological preons.}
Bilson-Thompson's braid model~\citep{BilsonThompson2005} maps the
first-generation Standard Model onto ribbon braids. RA is in the same
direction but derives the stability selection rule (the BDG filter)
that the braid model lacks. Whether RA's stable patterns correspond to
Bilson-Thompson braids is an open question that D1 would resolve.

% ─────────────────────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}
% ─────────────────────────────────────────────────────────────────────────────

Physics has spent a century seeking the reconciliation of quantum
mechanics and general relativity, and asking why the Standard Model
has the structure it has. Relational Actualism proposes that both
questions are symptoms of a shared error: treating spacetime,
matter, and quantum states as primitive, when none of them is.

The primitive is the actualization event. Not the particle, not
the field, not the wave function, not the spacetime manifold.
The moment at which a physical interaction becomes irreversible ---
at which $\Delta S(\rho\|\sigma_0) > 0$ and a permanent causal edge
is written into the DAG.

From this:

\medskip
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{ll}
\toprule
Result & Status \\
\midrule
$c = l_P/t_P$; $E = \gamma mc^2$; $\tau = N_{\mathrm{act}}\cdot t_P$ & Proved \\
GR inevitable (Lovelock) & Proved \\
Three non-gravitational forces in 4D & Proved \\
Charge quantisation in units of $e/3$ & Proved \\
Photon masslessness & Proved \\
Colour confinement & Proved \\
Exact baryon conservation & Proved (Lean-backed) \\
Maximal parity violation & Proved \\
Exactly three generations & Proved \\
Koide lepton mass formula & Conjectured; $K = 2/3$ exact \\
GUT unification at Planck scale & Conjecture \\
Flat rotation curves; Hubble tension & Derived \\
$\kappa(\theta)$ Bullet Cluster profile & Qualitative (RADM §4.2) \\
4-regime lensing classification & Established (RADM §4.3) \\
DJW lensing $\kappa(b) = v_\mathrm{flat}^2/(4G\Sigma_\mathrm{cr}b)$ & Derived (RADM §4.1) \\
$\kappa(b)$ cutoff at $R_\mathrm{baryon}$ vs NFW & Falsifiable prediction \\
Galaxy-galaxy amplitude (CGM) & Prediction, open calc \\
Origin of life (Causal Firewall) & Derived \\
QFT fault-tolerance threshold & Derived \\
Measurement problem & Dissolved \\
$\theta_{\mathrm{QCD}} = 0$ exactly; no axion & Proved \\
Black hole information paradox & Dissolved \\
Baryon asymmetry as Causal Severance IC & Derived \\
$d = 4$ spacetime dimensionality & Conjecture \\
Quark Koide breaking (RA form; $\alpha_s$ external input) & Conjecture \\
Wyler formula for $\alpha_{\mathrm{EM}}$ (0.00006\%) & Conjecture \\
$\theta_{\mathrm{lep}} = 2/9$ (Koide fit, $1:10^5$) & Conjecture \\
$|V_{us}| = \sin(2/9)$ Cabibbo prediction (1.8\%) & Conjecture \\
$\delta = \tfrac{3}{4}\alpha_s/\pi$ ($0.1\%$) & Proved \\
$|V_{cb}| = (2/\pi)\delta$ ($0.4\%$) & Conjecture \\
\bottomrule
\end{tabular}
\end{center}

\medskip
The question driving the next phase of the programme is exact:
is the complete set of $\SBDG > 0$-stable self-similar patterns
in a 4D BDG-filtered Poisson-CSG precisely the first-generation
Standard Model particle spectrum? If yes --- if the BDG-filter MCMC
confirms that the stable patterns are exactly $\{e, u, d, \nu_e,
\gamma, g, W, Z, H\}$ --- then all of physics follows from the
single fact that some things happen and others do not.

% ─────────────────────────────────────────────────────────────────────────────
\bibliographystyle{plainnat}
\begin{thebibliography}{99}

\bibitem[Sandeman(2026RACL)]{Sandeman2026RACL}
J.~F. Sandeman.
\newblock {Relational Actualism: The Einstein Field Equations as the Unique
  Macroscopic Limit of the Actualization Graph (RACL)}.
\newblock \textit{Preprint}, 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19270020}{doi.org/10.5281/zenodo.19270020}.


\bibitem[Sandeman(2026RATM)]{Sandeman2026RATM}
J.~F. Sandeman.
\newblock {Relational Actualism: The Topology of Matter (RATM)}.
\newblock \textit{Preprint}, 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19265651}{doi.org/10.5281/zenodo.19265651}.

\bibitem[Sandeman(2026RAEB)]{Sandeman2026RAEB}
J.~F. Sandeman.
\newblock {Relational Actualism: The Engine of Becoming (RAEB)}.
\newblock \textit{Preprint}, 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19198120}{doi.org/10.5281/zenodo.19198120}.


\bibitem[Benincasa \& Dowker(2010)]{Benincasa2010}
D.~M.~T.~Benincasa and F.~Dowker.
\newblock The scalar curvature of a causal set.
\newblock \emph{Physical Review Letters}, 104(18):181301, 2010.
\newblock \href{https://doi.org/10.1103/PhysRevLett.104.181301}%
  {doi:10.1103/PhysRevLett.104.181301}.

\bibitem[Bilson-Thompson(2005)]{BilsonThompson2005}
S.~O.~Bilson-Thompson.
\newblock A topological model of composite preons.
\newblock \emph{arXiv preprint}, 2005.
\newblock \href{https://arxiv.org/abs/hep-ph/0503213}{arXiv:hep-ph/0503213}.

\bibitem[Bombelli et~al.(1987)]{Bombelli1987}
L.~Bombelli, J.~Lee, D.~Meyer, and R.~D.~Sorkin.
\newblock Space-time as a causal set.
\newblock \emph{Physical Review Letters}, 59(5):521--524, 1987.
\newblock \href{https://doi.org/10.1103/PhysRevLett.59.521}%
  {doi:10.1103/PhysRevLett.59.521}.

\bibitem[Bose et~al.(2017)]{Bose2017}
S.~Bose et~al.
\newblock Spin entanglement witness for quantum gravity.
\newblock \emph{Physical Review Letters}, 119(24):240401, 2017.
\newblock \href{https://doi.org/10.1103/PhysRevLett.119.240401}%
  {doi:10.1103/PhysRevLett.119.240401}.

\bibitem[Koide(1983)]{Koide1983}
Y.~Koide.
\newblock New view of quark and lepton mass hierarchy.
\newblock \emph{Physical Review D}, 28(1):252--254, 1983.
\newblock \href{https://doi.org/10.1103/PhysRevD.28.252}%
  {doi:10.1103/PhysRevD.28.252}.

\bibitem[Lovelock(1971)]{Lovelock1971}
D.~Lovelock.
\newblock The Einstein tensor and its generalizations.
\newblock \emph{Journal of Mathematical Physics}, 12(3):498--501, 1971.
\newblock \href{https://doi.org/10.1063/1.1665613}{doi:10.1063/1.1665613}.

\bibitem[Marletto \& Vedral(2017)]{Marletto2017}
C.~Marletto and V.~Vedral.
\newblock Gravitationally induced entanglement between two massive particles
  is sufficient evidence of quantum effects in gravity.
\newblock \emph{Physical Review Letters}, 119(24):240402, 2017.
\newblock \href{https://doi.org/10.1103/PhysRevLett.119.240402}%
  {doi:10.1103/PhysRevLett.119.240402}.

\bibitem[Particle Data Group(2022)]{PDG2022}
R.~L.~Workman et~al.
\newblock Review of particle physics.
\newblock \emph{Progress of Theoretical and Experimental Physics},
  2022(8):083C01, 2022.
\newblock \href{https://doi.org/10.1093/ptep/ptac097}{doi:10.1093/ptep/ptac097}.

\bibitem[Sandeman(2026a)]{Sandeman2026RAQM}
J.~F.~Sandeman.
\newblock Relational Actualism and Quantum Mechanics (RAQM).
\newblock \emph{Foundations of Physics} (under review), 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19174380}%
  {doi:10.5281/zenodo.19174380}.

\bibitem[Sandeman(2026b)]{Sandeman2026RAGC}
J.~F.~Sandeman.
\newblock Relational Actualism: Gravity and Cosmology (RAGC).
\newblock Preprint, 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19197999}%
  {doi:10.5281/zenodo.19197999}.

\bibitem[Sandeman(2026c)]{Sandeman2026EB}
J.~F.~Sandeman.
\newblock The Engine of Becoming (EB).
\newblock Preprint, 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19198120}%
  {doi:10.5281/zenodo.19198120}.

\bibitem[Sandeman(2026d)]{Sandeman2026RAHC}
J.~F.~Sandeman.
\newblock Relational Actualism: Algorithmic Causal Dynamics (RAHC).
\newblock Preprint, 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19198171}%
  {doi:10.5281/zenodo.19198171}.

\bibitem[Sandeman(2026e)]{Sandeman2026RACI}
J.~F.~Sandeman.
\newblock Relational Actualism: Complexity and Intelligence (RACI).
\newblock Preprint, 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19198224}%
  {doi:10.5281/zenodo.19198224}.

\bibitem[Sandeman(2026f)]{Sandeman2026RAQI}
J.~F.~Sandeman.
\newblock Relational Actualism: Implications for Quantum Information (RAQI).
\newblock Preprint, 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19198285}%
  {doi:10.5281/zenodo.19198285}.

\bibitem[Sandeman(2026g)]{Sandeman2026RADM}
J.~F.~Sandeman.
\newblock Relational Actualism: Dark Matter as Actualization-Density
  Topology (RADM).
\newblock Preprint, 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19198513}%
  {doi:10.5281/zenodo.19198513}.

\bibitem[Glaser(2014)]{Glaser2014}
L.~Glaser.
\newblock A closed form expression for the causal set d'Alembertian.
\newblock \emph{Journal of Mathematical Physics}, 55(9):092301, 2014.
\newblock \href{https://arxiv.org/abs/1311.1701}{arXiv:1311.1701}.

\bibitem[Page(1993)]{Page1993}
D.~N.~Page.
\newblock Information in black hole radiation.
\newblock \emph{Physical Review Letters}, 71(23):3743--3746, 1993.
\newblock \href{https://doi.org/10.1103/PhysRevLett.71.3743}{doi:10.1103/PhysRevLett.71.3743}.

\bibitem[Sandeman(2026h)]{Sandeman2026RASM}
J.~F.~Sandeman.
\newblock Relational Actualism and the Standard Model (RASM).
\newblock Preprint, 2026.
\newblock \href{https://doi.org/10.5281/zenodo.19229335}{doi.org/10.5281/zenodo.19229335}

\bibitem[Hawking(1975)]{Hawking1975}
S.~W.~Hawking.
\newblock Particle creation by black holes.
\newblock \emph{Communications in Mathematical Physics}, 43(3):199--220, 1975.
\newblock \href{https://doi.org/10.1007/BF02345020}{doi:10.1007/BF02345020}.

\bibitem[Particle Data Group(2022)]{PDG2022}
R.~L.~Workman et~al.
\newblock Review of particle physics.
\newblock \emph{Progress of Theoretical and Experimental Physics},
  2022(8):083C01, 2022.
\newblock \href{https://doi.org/10.1093/ptep/ptac097}{doi:10.1093/ptep/ptac097}.

\bibitem[Peccei \& Quinn(1977)]{Peccei1977}
R.~D.~Peccei and H.~R.~Quinn.
\newblock CP conservation in the presence of pseudoparticles.
\newblock \emph{Physical Review Letters}, 38(25):1440--1443, 1977.
\newblock \href{https://doi.org/10.1103/PhysRevLett.38.1440}{doi:10.1103/PhysRevLett.38.1440}.

\bibitem[Sakharov(1967)]{Sakharov1967}
A.~D.~Sakharov.
\newblock Violation of CP invariance, C asymmetry, and baryon asymmetry of the universe.
\newblock \emph{JETP Letters}, 5:24--27, 1967.

\bibitem[Sorkin(1994)]{Sorkin1994}
R.~D.~Sorkin.
\newblock Quantum mechanics as quantum measure theory.
\newblock \emph{Modern Physics Letters A}, 9(33):3119--3127, 1994.
\newblock \href{https://doi.org/10.1142/S021773239400294X}%
  {doi:10.1142/S021773239400294X}.

\bibitem[Wyler(1969)]{Wyler1969}
A.~Wyler.
\newblock On the conformal groups in the theory of relativity and their
  unitary representations.
\newblock \emph{Comptes Rendus Acad.~Sci.~Paris}, 269:743--745, 1969.

\bibitem[Walker et~al.(2023)]{Walker2023}
S.~I.~Walker et~al.
\newblock Explaining the origin of life is harder than we think.
\newblock \emph{arXiv preprint}, 2023.
\newblock \href{https://arxiv.org/abs/2308.09738}{arXiv:2308.09738}.

\end{thebibliography}

\end{document}
